---
title: "Case Study"
author: "Isiah Perine"
date: "2024-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```



# LOADING IN THE DATA

```{r}
wine = read.csv("Wine Train.csv")
wine
```



# CHECKING NA'S

```{r}
sum(is.na(wine))
```

```{r}
# Example dataset
data <- data.frame(old_name = 1:5, another_var = 6:10)

# View original names
names(data)

# Rename a specific variable
names(data)[names(data) == "old_name"] <- "new_name"

# View updated names
names(data)

```


# R SHINY APP FOR SCATTERPLOT/BOXPLOT OF EVERY VARIABLE AGAINST QUALITY

```{r}
library(shiny)
library(ggplot2)

data = read.csv("Wine Train.csv")
names(data)[names(data) == "fixed.acidity"] = "FixedAcidity" 
names(data)[names(data) == "volatile.acidity"] = "VolatileAcidity"
names(data)[names(data) == "citric.acid"] = "CitricAcid"
names(data)[names(data) == "residual.sugar"] = "ResidualSugar"
names(data)[names(data) == "chlorides"] = "Chlorides"
names(data)[names(data) == "free.sulfur.dioxide"] = "FreeSulfurDioxide"
names(data)[names(data) == "total.sulfur.dioxide"] = "TotalSulfurDioxide"
names(data)[names(data) == "density"] = "Density"
names(data)[names(data) == "sulphates"] = "Sulphates"
names(data)[names(data) == "alcohol"] = "Alcohol"
names(data)[names(data) == "quality"] = "Quality"

ui = fluidPage(
    titlePanel("Scatterplot and Boxplot of Variables vs. Quality"),
    sidebarLayout(
        sidebarPanel(
            selectInput("variable", 
                        "Select a Variable:", 
                        choices = setdiff(names(data), "Quality"), 
                        selected = names(data)[1]),
            radioButtons("plotType", 
                         "Select Plot Type:", 
                         choices = c("Scatterplot" = "scatter", "Boxplot" = "box"),
                         inline = TRUE)
        ),
        mainPanel(
            plotOutput("variablePlot")
        )
    )
)

server = function(input, output) {
    output$variablePlot = renderPlot({
        if (input$plotType == "scatter") {
            ggplot(data, aes_string(x = "Quality", y = input$variable)) +
                geom_point(alpha = 0.7, color = "blue") +
                labs(title = paste("Scatterplot of", input$variable, "vs Quality"),
                     x = "Quality",
                     y = input$variable) +
                theme_minimal()
        } else {
            ggplot(data, aes(x = factor(Quality), y = .data[[input$variable]], fill = factor(Quality))) +
                geom_boxplot() +
                labs(title = paste("Boxplot of", input$variable, "by Quality"),
                     x = "Quality",
                     y = input$variable) +
                scale_fill_brewer(palette = "Set3") +
                theme_minimal() +
                theme(legend.position = "none")
        }
    })
}

shinyApp(ui = ui, server = server)
```


```{r}
wine

ggplot(wine, aes(x = factor(quality), y = volatile.acidity, fill = factor(quality))) +
  geom_boxplot() +
  labs(title = "Boxplot of Volatile Acidity by Quality",
       x = "Quality",
       y = "Volatile Acidity") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
library(shiny)
library(ggplot2)

data = read.csv("Wine Train.csv")

ui = fluidPage(
    titlePanel("Scatterplot and Boxplot of Variables vs. Quality"),
    sidebarLayout(
        sidebarPanel(
            selectInput("variable", 
                        "Select a Variable:", 
                        choices = setdiff(names(data), "quality"), 
                        selected = names(data)[1]),
            radioButtons("plotType", 
                         "Select Plot Type:", 
                         choices = c("Scatterplot" = "scatter", "Boxplot" = "box"),
                         inline = TRUE)
        ),
        mainPanel(
            plotOutput("variablePlot")
        )
    )
)

server = function(input, output) {
    output$variablePlot = renderPlot({
        if (input$plotType == "scatter") {
            ggplot(data, aes_string(x = "quality", y = input$variable)) +
                geom_point(alpha = 0.7, color = "blue") +
                labs(title = paste("Scatterplot of", input$variable, "vs Quality"),
                     x = "Quality",
                     y = input$variable) +
                theme_minimal()
        } else {
            ggplot(data, aes(x = factor(quality), y = .data[[input$variable]], fill = factor(quality))) +
                geom_boxplot() +
                labs(title = paste("Boxplot of", input$variable, "by Quality"),
                     x = "Quality",
                     y = input$variable) +
                scale_fill_brewer(palette = "Set3") +
                theme_minimal()
        }
    })
}

shinyApp(ui = ui, server = server)
```

- As the quality of wine increases, the distribution of citirc.acid increases as well
- After a quality level of 5, as the quality of wine increases, the distribution of density decreases
- After a quality level of 5, as the quality of wine increases, the distribution of alcohol increases as well



# CALCULATING ALL CORRELATIONS

```{r}
correlations = sapply(wine, function(x) {
  if (is.numeric(x) && is.numeric(wine$quality)) {
    cor(wine$quality, x, use = "complete.obs")
  } else {
    NA
  }
})

correlation_results = data.frame(Variable = names(correlations), Correlation = correlations)

# Sort the data by 'Correlation' in ascending order
sorted_correlations = correlation_results[order(correlation_results$Correlation), ]
sorted_correlations
```

- alcohol
- density
- volatile.acidity
- chlorides
- citric.acid



# R SHINY APP FOR RELATIONSHIP OF EVERY VARIABLE AGAINST EACH OTHER COLORED BY GOOD/BAD QUALITY

```{r}
library(shiny)
library(ggplot2)

data = read.csv("Wine Train.csv")

data$goodbad = ifelse(data$quality >= 7, "good", "bad")

ui = fluidPage(
    titlePanel("Dynamic Plot with Good/Bad Classification"),
    sidebarLayout(
        sidebarPanel(
            selectInput("xvar", "Select X Variable:", choices = names(data)),
            selectInput("yvar", "Select Y Variable:", choices = names(data), selected = names(data)[2])
        ),
        mainPanel(
            plotOutput("dynamicPlot")
        )
    )
)

server = function(input, output) {
    output$dynamicPlot = renderPlot({
        ggplot(data, aes_string(x = input$xvar, y = input$yvar, color = "goodbad")) +
            geom_point(alpha = 0.7) +
            labs(title = paste("Plot of", input$xvar, "vs", input$yvar),
                 x = input$xvar, y = input$yvar, color = "Good/Bad") +
            theme_minimal()
    })
}

shinyApp(ui = ui, server = server)
```



# R SHINY APP FOR BOXPLOT OF EVERY VARIABLE AGAINST GOOD/BAD QUALITY

```{r}
library(shiny)
library(ggplot2)

data = read.csv("Wine Train.csv")

data$goodbad = ifelse(data$quality >= 7, "good", "bad")

ui = fluidPage(
    titlePanel("Boxplots for Variables by Good/Bad Classification"),
    sidebarLayout(
        sidebarPanel(
            selectInput("variable", "Select a Variable for Boxplot:", choices = names(data))
        ),
        mainPanel(
            plotOutput("boxplot")
        )
    )
)

server = function(input, output) {
    output$boxplot = renderPlot({
        ggplot(data, aes(x = goodbad, y = .data[[input$variable]], fill = goodbad)) +
            geom_boxplot() +
            labs(
                title = paste("Boxplot of", input$variable, "by Good/Bad Classification"),
                x = "Good/Bad Classification",
                y = input$variable
            ) +
            scale_fill_manual(values = c("good" = "lightgreen", "bad" = "tomato")) +
            theme_minimal()
    })
}

shinyApp(ui = ui, server = server)
```

- The distribution of the alcohol is higher for the good quality wine compared to the bad quality wine. 



```{r}

```

